\part{Creating Full Documents in \LaTeX}
\section{Starting Point}
\begin{itemize}
\item This tutorial builds off of the previous one which introduced
  \LaTeX~commands.
\item By the end of the last tutorial we had working documents with
  environments and headings that were quite pleasing to the eye.
\item We will quickly create a very basic document to pick up where we
  left off.
\item Open WinEdt and create a new \LaTeX~\texttt{*.tex} file where
  you would like it.
\item Create a basic document with the following code and check that
  it compiles.
  
  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}
\documentclass{article}

% This is just a comment.
\begin{document}

\section{The Content}

\subsection{Short Content}

\subsection{Long Content}

\appendix

\section{Some Book-keeping}

\end{document}
\end{verbatim}
    \end{minipage}
  }
\item Go ahead and download the two dummy text files from
  \url{www.rochester.edu/college/gradstudents/jolmsted/teaching/LaTeX/}. Place
  these two files inside the same directory as your new
  \LaTeX~\texttt{.tex} file.
\item As it stands, our compiled document is quite bare. Add the
  following \LaTeX~commands:
  \begin{itemize}
  \item Add \texttt{\textbackslash input\{./loremipsum\_short.txt\}}
    on a new line directly after \texttt{\textbackslash section\{Short
      Content\}}. As long as the downloaded file was placed in the
    appropriate place, you can safely ignore what this command does.
  \item Add \texttt{\textbackslash input\{./loremipsum.txt\}} on a new
    line directly after \texttt{\textbackslash
      section\{Long~Content\}}. Enter this line, again, on a new
    line. As long as the downloaded file was placed in the appropriate
    place, you can safely ignore what this command does.
  \end{itemize}
\item Compile this document again.
\end{itemize}

\section{Fonts}
In \LaTeX~we can make some very basic changes to the font in with
ease. These changes can be made with one of two approaches. The
\textit{environment} approach or the \textit{command} approach. In the
first, whatever font we change will be inside an environment. In the
second, whatever font we change will be the mandatory argument to a
command.

\subsection*{Font Styles}
When consider \textbf{bold face} or \textsc{Small Caps}, we are
changing font styles. The basic and most common changes are here:
\begin{small}
\begin{center}
  \begin{tabular}{c c c}
    \hline
    Style               & Environment                                                                          & Command \\
    \hline
    normal/default      & \texttt{\textbackslash begin\{textnormal\} \ldots \textbackslash end\{textnormal\}}  & \texttt{\textbackslash textnormal\{\ldots \}}\\
   \textbf{bold}        & \texttt{\textbackslash begin\{bf\} \ldots \textbackslash end\{bf\}}  & \texttt{\textbackslash textbf\{\ldots \}}\\
   \textit{italics}     & \texttt{\textbackslash begin\{it\} \ldots \textbackslash end\{it\}}  & \texttt{\textbackslash textit\{\ldots \}}\\
   \textsc{SmallCaps}   & \texttt{\textbackslash begin\{sc\} \ldots \textbackslash end\{sc\}}  & \texttt{\textbackslash textsc\{\ldots \}}\\
   \underline{underline}&                                      :-(                             & \texttt{\textbackslash underline\{\ldots \}}\\
   \textrm{Roman}       & \texttt{\textbackslash begin\{textrm\} \ldots \textbackslash end\{textrm\}}  & \texttt{\textbackslash textrm\{\ldots \}}\\
   \textsf{Sans Serif}  & \texttt{\textbackslash begin\{textsf\} \ldots \textbackslash end\{textsf\}}  & \texttt{\textbackslash textsf\{\ldots \}}\\
   \texttt{teletype}    & \texttt{\textbackslash begin\{tt\} \ldots \textbackslash end\{tt\}}  & \texttt{\textbackslash texttt\{\ldots \}}\\
   \hline
  \end{tabular}
\end{center}
\end{small}
\begin{itemize}
\item Place the \texttt{\textbackslash input} command under the ``Short Content''
  section inside a font style environment of your choice. Compile the
  document.
\item Before this environment, but under the ``Short Content'' section
  type and complete the following sentence: ``If I had a boat, I'd \ldots''.
\item Change the style of the font for that sentence and only that
  sentence by using the command approach. This time, choose a
  different style from the environment change you made above.
\end{itemize}



\subsection*{Sizes}

\begin{small}
\begin{center}
  \begin{tabular}{c c c}
    \hline
    Style               & Environment                                                                 & Command \\
    \hline
    \tiny{tiny}         & \texttt{\textbackslash begin\{tiny\} \ldots \textbackslash end\{tiny\}}      & \texttt{\textbackslash tiny\{\ldots\}}\\
    \scriptsize{scriptsize} & \texttt{\textbackslash begin\{scriptsize\} \ldots \textbackslash end\{scriptsize\}}      & \texttt{\textbackslash scriptsize\{\ldots\}}\\
    \footnotesize{footnotesize} & \texttt{\textbackslash begin\{footnotesize\} \ldots \textbackslash end\{footnotesize\}}      & \texttt{\textbackslash footnotesize\{\ldots\}}\\
    \small{small} & \texttt{\textbackslash begin\{small\} \ldots \textbackslash end\{small\}}      & \texttt{\textbackslash small\{\ldots\}}\\
    \normalsize{normalsize} & \texttt{\textbackslash begin\{normalsize\} \ldots \textbackslash end\{normalsize\}}      & \texttt{\textbackslash normalsize\{\ldots\}}\\
    \large{large} & \texttt{\textbackslash begin\{large\} \ldots \textbackslash end\{large\}}      & \texttt{\textbackslash large\{\ldots\}}\\
    \Large{Large} & \texttt{\textbackslash begin\{Large\} \ldots \textbackslash end\{Large\}}      & \texttt{\textbackslash Large\{\ldots\}}\\
    \LARGE{LARGE} & \texttt{\textbackslash begin\{LARGE\} \ldots \textbackslash end\{LARGE\}}      & \texttt{\textbackslash LARGE\{\ldots\}}\\
    \huge{huge} & \texttt{\textbackslash begin\{huge\} \ldots \textbackslash end\{huge\}}      & \texttt{\textbackslash huge\{\ldots\}}\\
    \Huge{Huge} & \texttt{\textbackslash begin\{Huge\} \ldots \textbackslash end\{Huge\}}      & \texttt{\textbackslash Huge\{\ldots\}}\\
    \hline
  \end{tabular}
\end{center}
\end{small}

\begin{itemize}
\item Place the \texttt{\textbackslash input} under the ``Long
  Content'' inside a font size environment of your choice, but choose
  a size that is small than the normal size. Compile your document.
\end{itemize}

\section{List Environments}
There are three main list environments that you will use:
\texttt{itemize}, \texttt{enumerate}, and \texttt{description}.

\begin{description}
\item[\texttt{itemize}] The \texttt{itemize} environment allows you to
  make bullet-ed lists without conveying any order. These can be
  nested to create any structure desired. However, complex nesting
  often results in user-mistakes.
\item[\texttt{enumerate}] The \texttt{enumerate} environment allows
  you to make numbered lists which convey a precise order. The number
  is automatic. These can be nested to create any structure desired
  and the ``numbering'' system used at each level reflects how far in
  the tree the entry is.
\item[\texttt{description}] The \texttt{description} environment is
  slightly different. In fact, I am using it here. Instead of each
  entry being marked with an automatic bullet or number, the user must
  provide the ``term'' being described.
\end{description}

\begin{itemize}
\item For either \texttt{itemize} or \texttt{enumerate} your code
  looks like this, where ``\texttt{environment}'' is swapped out for
  the correct environment:

  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}
\begin{environment}
\item An item
\item Another item
\item Still more items!
\end{environment}
\end{verbatim}
    \end{minipage}
  }  


\item For the \texttt{description} environment your code is slightly
  different:

  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}
\begin{description}
\item[An item] description 1
\item[Another item] description 2
\item[Still more items!] description 3
\end{description}
\end{verbatim}
    \end{minipage}
  }  

\item Create a list in the ``Some Book-keeping'' section. In this list
  you should have two items---two of your favorite books. If you are
  using the \texttt{description} environment include the author's name
  as the description if you know it. In this case, make sure the book
  title is in square brackets.
\end{itemize}

\section{Metadata: titles, dates, authors, abstracts}
\subsection*{\texttt{\textbackslash maketitle}}
\begin{itemize}
\item Although our content is already impressively long, the document
  is lacking all metadata. To address this we will formally introduce
  the \texttt{\textbackslash maketitle} command
\item In the preamble of the file (that is, before
  \texttt{\textbackslash begin\{document\}}, but after
  \texttt{\textbackslash documentclass}), enter the following code:

  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}
\title{My First Opus}
\date{\today}
\author{Kevin Clarke}
\end{verbatim}
    \end{minipage}
  }
\item Compile the document. Does anything change?
\item Now, add \texttt{\textbackslash maketitle} directly after
  \texttt{\textbackslash begin\{document\}}, but on a new line.
\item Compile again.
\item Comment out the line with the \texttt{\textbackslash date}
  directive. Re-compile. What happens?
\item Uncomment that same line, delete the directive's single
  mandatory argument and compile. What happens?
\item In order to use the \texttt{\textbackslash maketitle} command, a
  title must be defined. The author and date, however are optional.
\end{itemize}

\subsection*{\texttt{\textbackslash thanks}}
\begin{itemize}
\item Kevin is a humble man and acknowledges the contribution that
  non-Kevin entities play in his success. To add this to the title we
  use \texttt{\textbackslash thanks}.
\item Include the following line at the end of ``Clarke'' within the
  author directive as part of the mandatory
  argument: \texttt{\textbackslash thanks\{The author thanks Rochester
    winters for being sufficiently ridiculous that everyone is forced
    to stay inside and work.\}}
\item You can use \texttt{thanks} to include institution information
  or contact information, too.
\end{itemize}



\subsection*{\texttt{titlepage}}
\begin{itemize}
\item Because this opus is so substantial, it can be overwhelming to
  assault the reader with content on the first page. In this case,
  we'd like to have just a title page.
\item Place the \texttt{\textbackslash maketitle} directive in an
  environment by the name of \texttt{titlepage}
\end{itemize}

\subsection*{\texttt{abstract}}
\begin{itemize}
\item Even with a title, we might want to convey the content of our
  document in more detail.
\item In this case, we'd use an abstract.
\item After \texttt{\textbackslash maketitle} but within the
  \texttt{titlepage} environment, create a new environment (which is
  completely nested) called \texttt{abstract}. Within the
  \texttt{abstract} environment, write two sentences about what we
  have covered in this tutorial, so far. Include the word \LaTeX,
  properly typeset.
\item Compile the document.
\end{itemize}

\section{Fancy (and not Fancy) Headers}

\begin{itemize}
\item By default, the only content in the header or footer of our
  document is the page number.
\item We can change this by using a package called \texttt{fancyhdr}.
\item Tell \LaTeX~ to use this package? What is the command? Where
  does it go?
\item In the header, also include \texttt{\textbackslash
    pagestyle\{plain\}}. Compile the document. What changes?
\item Now use \texttt{\textbackslash pagestyle\{fancy\}},
  instead. Compile the document. What changes?
\item You can change the contents of the headers and footers by using
  commands of the form \texttt{\textbackslash rfoot\{\}} for
  \texttt{r}, \texttt{l}, and \texttt{c}; and for \texttt{foot} and
  \texttt{head}.
\item Place the state or non-US country in which you were born
  somewhere in the footer. Compile the document.
\item Notice that the first page, the title page, has no header. Even
  when we use fancy headers and get the pretty content on that first
  inch of the page, any page with \texttt{\textbackslash maketitle} is
  unchanged, by default. We can change this by including
  \texttt{\textbackslash thispagestyle\{fancy\}} immediately after
  \texttt{\textbackslash maketitle}.
\end{itemize}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 


%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
