\part{Including Mathematics in \LaTeX{} Documents}

\section{Additional Packages}

\begin{itemize}

\item In order to avoid a lot of redundant and complicated code, it is
  helpful to load two particular packages each time you create a
  document with mathematics in it.

\item A bare-bones document requires only a \texttt{\textbackslash
    documentclass\{\}} line, followed by the \texttt{\textbackslash
    begin\{document\}} and \texttt{\textbackslash
    end\{document\}} lines.

\item We will now add \texttt{\textbackslash usepackage\{amsmath,
    amsthm, amssym, amsfonts\}} to the \textit{preamble} of the
  document. This tells \LaTeX{} to recognize certain commands defined
  in these packages.

\item We'll never use most of the functionality they add, but we'll
  often use some of it!

\item Create the following empty \LaTeX{} document in WinEdt in which
  we'll practice mathematical commands.

  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}
\documentclass{article}

\usepackages{amsmath, amsthm, amssymb, amsfonts}

% This is just a comment.

\begin{document}

\end{document}
\end{verbatim}
    \end{minipage}
  }

\end{itemize}

\section{Math Modes}

\begin{itemize}
  
\item In order to display mathematical content in \LaTeX{}, we will use a
  new set of commands that we wouldn't need for, say, writing a
  letter. However, \LaTeX{} doesn't allow you to issue these commands
  just anywhere. 

\item Rather, you need to be either in math mode or in a
  mathematical \textit{environment} where the math mode is implied.

\item There are two math modes which don't utilize special
  \textit{environments}: \textit{in-line} math and \textit{display
    math}.

\item These two are analogous to the way we include quotes in our own
  written work. \subparagraph{In-line} Sometimes the quotes are
  treated like any other text and printed in the same sized font
  without any special indentation. \subparagraph{Display} Sometimes
  the quotes begin on new lines, are indented, and quite distinct from
  the other text.

\item As a note, some commands and characters can only be entered in
  math mode. However, many, like the word ``computer'' can be entered
  in normal mode and math mode. Yet, if you think ``computer'' is
  different from ``$computer$'', then you must beware which mode you
  use.

\item After this section, it will be assumed that all math command
  sequence are typed in math mode. Otherwise, they will almost always
  fail. The math mode openings and closings are omitted.
  
\end{itemize}

\subsection*{In-line Math}
\begin{itemize}
\item The \textit{in-line} math mode is opened by typing a single
  dollar sign (\texttt{\$}) and then closed by typing a single dollar
  sign.
\item The content is placed in between the two \texttt{\$} symbols.
\item When we type \texttt{\$55/i = \textbackslash pi\^{}\{-0.3\}\$},
  we get $55/i = \pi^{-0.3}$. Any lines immediately following the
  in-line text are entirely unaffected, including this one. And
  this one, too.
\end{itemize}

\subsection*{Display Math}
\begin{itemize}
\item The \textit{display} math mode is opened by typing
  \texttt{\textbackslash [} and closed by typing
  \texttt{\textbackslash ]}. Another approach is to use double dollar
  signs as in the in-line mode (\texttt{\$\$ \ldots \$\$}). However,
  the square bracket approach is considered ``better''.

\item Again, the content to be displayed in the math mode is placed
  between the opening and closing key sequences.
\item Now, when we type \texttt{\textbackslash [55/i = \textbackslash
    pi\^{}\{-0.3\}\textbackslash ]}, we get
  \[
  55/i = \pi^{-0.3},
  \]
  which is the same as when we type \texttt{\$\$55/i = \textbackslash
    pi\^{}\{-0.3\}\$\$} and we get
  $$
  55/i = \pi^{-0.3}.
  $$
\item Clearly the appearance of the \textit{display} math is different
  from the \textit{in-line} math. However, sometimes the differences
  go beyond indentation and size. Both $\int_{0}^{\infty}
  \sum_{n=1}^{\infty} \frac{1}{n}~dx$ and 
\[
\int_{0}^{\infty} \sum_{n=1}^{\infty} \frac{1}{n}~dx
\]
have the same math mode code, but they are placed in the two different
modes. In this examples, the limits of integration are the most
obvious difference in how these expressions are rendered.
\end{itemize}

\section{Superscripts and Subscripts}
\begin{itemize}
\item To add a superscript to an expression use a hat,
  ``\texttt{\^{}}'', after the term with the exponent. So,
  \texttt{X\^{}3} gives $X^3$.

\item If your exponent has more than one character to be rendered in
  the case of $X^{3+y}$, then we must surround the entire exponent in
  \textit{curly braces} (\texttt{\{ \}}). For this previous expression
  we'd type \texttt{X\^{}\{3+y\}}. In this sense, the
  \textit{curly braces} group terms. When in doubt, use curly braces
  to set the scope of your exponent. Had we omitted the curly braces,
  we'd see $X^3+y$ which is quite different.

\item The principle for subscripting is identical, but we use the
  underscore, ``\texttt{\_}'', instead. Typing
  \texttt{X\_{}\{3+y\}} gives $X_{3 + y}$.

\item Super-- and subscripting can be recursive. However, one must be
  very careful with braces in this case. We can typeset $e^{x_{i}}$
  with \texttt{e\^{}\{x\_\{i\}\}}. Now, the braces around the
  \texttt{i} term are optional. Remove them, what happens to the
  rendering? Remove the braces around the $x_i$ term. Compile the
  document. What happens to the rendering?

\item Both superscripts and subscripts can be used simultaneously. In
  Social Choice, we often care about the median voter's preference,
  denoted $x_i^m$. This can be achieved with either
  \texttt{x\_i\^{}m} or \texttt{x\_\{i\}\^{}\{m\}}. The latter
  is more typing, but forces you to be explicit about the scope which
  tends to reduce the number of errors.

\end{itemize}

\section{Operators}

\begin{multicols}{3}
\begin{itemize}
\item[$+$] \texttt{+}
\item[$=$] \texttt{=}
\item[$-$] \texttt{-}
\item[$/$] \texttt{/}
\item[$<$] \texttt{<}
\item[$>$] \texttt{>}
\item[$\leq$] \texttt{\textbackslash leq}
\item[$\geq $] \texttt{\textbackslash geq}
\item[$\vee$] \texttt{\textbackslash vee}
\item[$\wedge$] \texttt{\textbackslash wedge}
\item[$\bigvee$] \texttt{\textbackslash bigvee}
\item[$\bigwedge$] \texttt{\textbackslash bigwedge}
\item[$\cup$] \texttt{\textbackslash cup}
\item[$\cap$] \texttt{\textbackslash cap}
\item[$\bigcup$] \texttt{\textbackslash bigcup}
\item[$\bigcap$] \texttt{\textbackslash bigcap}
\item[$\forall$] \texttt{\textbackslash forall}
\item[$\exists$] \texttt{\textbackslash exists}
\item[$\subset$] \texttt{\textbackslash subset}
\item[$\subseteq$] \texttt{\textbackslash subseteq}
\item[$\subsetneq$] \texttt{\textbackslash subsetneq}
\item[$\supset$] \texttt{\textbackslash supset}
\item[$\supseteq$] \texttt{\textbackslash supseteq}
\item[$\supsetneq$] \texttt{\textbackslash supsetneq}
\item[$\int$] \texttt{\textbackslash int}
\item[$\sum$] \texttt{\textbackslash sum}
\item[$\prod$] \texttt{\textbackslash prod}
\item[$\partial$] \texttt{\textbackslash partial}
\item[$\sim$] \texttt{\textbackslash sim}
\item[$\approx$] \texttt{\textbackslash approx}
\item[$\Leftrightarrow$] \texttt{\textbackslash Leftrightarrow}
\item[$\leftrightarrow$] \texttt{\textbackslash leftrightarrow}
\item[$\Leftarrow$] \texttt{\textbackslash Leftarrow}
\item[$\leftarrow$] \texttt{\textbackslash leftrightarrow}
\item[$\Rightarrow$] \texttt{\textbackslash Rightarrow}
\item[$\rightarrow$] \texttt{\textbackslash rightarrow}
\item[$\not ~$] \texttt{\textbackslash not}
\item[$\neg$] \texttt{\textbackslash neg}
\item[$\div$] \texttt{\textbackslash div}
\item[$\infty$] \texttt{\textbackslash infty}
\item[$\emptyset$] \texttt{\textbackslash emptyset}
\item[$\in$] \texttt{\textbackslash in}
\end{itemize}
\end{multicols}

\section{Greek Letters}

\begin{itemize}
\item Greek letters are uncomfortably common in the work we do, so you
  will get used to including them in your documents.

\item Although there are neither $\zeta$ nor $\chi$ keys on your
  keyboard, we can include these symbols in the math environment in a
  very intuitive way. Type \texttt{\textbackslash name-of-letter}
  or \texttt{\textbackslash Name-of-letter} for the lower-case and
  upper-case version of the Greek letter.

\item So, \texttt{\textbackslash delta} and \texttt{\textbackslash
    Delta} give $\delta$ and $\Delta$, respectively.

\item Not every letter has a special upper-case version
  provided. Although \texttt{\textbackslash beta} gives $\beta$
  \texttt{\textbackslash Beta} is undefined.

\item Not every letter has a special character, regardless of
  case. For example, the Greek omicron is just like the Roman `o', so
  \texttt{o} gives $o$ which will be sufficient.

\item With Greek letters a particular issue arises often, although it
  isn't specific to Greek letters. \LaTeX{} commands do not need a
  leading space before the \texttt{\textbackslash}. Yet, they do need
  a trailing space so that \LaTeX{} knows the name of the command is
  over. So, \texttt{X\textbackslash beta} ($X\beta$) is rendered
  identically to \texttt{X~\textbackslash beta} ($X \beta$)
  because \LaTeX{} has its own way of interpreting
  whitespace. However, \texttt{X\textbackslash beta + \textbackslash
    epsilon} ($X\beta + \epsilon$) works whereas
  \texttt{X~\textbackslash beta+ \textbackslash epsilon} will not
  because \texttt{\textbackslash beta+} is an undefined control
  sequence (i.e. we just made up that command).

\end{itemize}
\section{Other Letter-y Things}

Sometimes we need to present change the presentation of standard
letters or symbols in math mode. Here are several common examples. The
way these work is that whatever symbols are to have their face changed
are passed to these commands as arguments. Notice how the ``math''
faces remove white space

\begin{tabular}{c c c c}
  \hline \hline
  Modification & Command & Example & Common Uses\\
  \hline
  Normal Math Face & ---& $ABC XY$ & most maths\\
  Roman Face & \texttt{\textbackslash textrm\{\}} &  $\textrm{ABC XY}$& text within equation\\
  Bold Math Face & \texttt{\textbackslash mathbf\{\}} &  $\mathbf{ABC XY}$ & vectors and matrices\\
  Blackboard Math & \texttt{\textbackslash mathbb\{\}} &  $\mathbb{ABC XY}$ & special sets of numbers\\
  Calligraphic Math & \texttt{\textbackslash mathcal\{\}} & $\mathcal{ABC XY}$& arbitrary sets\\
  \hline \hline
\end{tabular}

\section{Special Text-like Mathematical Expressions}

Certain functions, operators, and constructs pop up in math frequently
which are basically abbreviates for words. The cosine function---as in
$\cos(\pi)=1$---and the limit of an expression---as in $\lim_{x
  \rightarrow \infty} \frac{1}{x}$---are two such examples. If we just
typed \texttt{\$cos\$} ($cos$) or \texttt{\$lim\$} ($lim$) we get
results quite different. An incomplete list of the expressions for
which the control sequences are defined is below.

\begin{multicols}{4}
\begin{itemize}
\item[$\det$] \texttt{\textbackslash det}
\item[$\lim$] \texttt{\textbackslash lim}
\item[$\max$] \texttt{\textbackslash max}
\item[$\min$] \texttt{\textbackslash min}
\item[$\inf$] \texttt{\textbackslash inf}
\item[$\sup$] \texttt{\textbackslash sup}
\item[$\cos$] \texttt{\textbackslash cos}
\item[$\sin$] \texttt{\textbackslash sin}
\item[$\tan$] \texttt{\textbackslash tan}
\item[$\exp$] \texttt{\textbackslash exp}
\item[$\Pr$] \texttt{\textbackslash Pr}
\item[$\arg$] \texttt{\textbackslash arg}
\end{itemize}
\end{multicols}

\section{Delimiters}
Many times, we need to express groupings through the use of
delimiters. For example, $(100 - 100)^{100}$ is quite different than
$100 - 100^{100}$. For parentheses we can just enter \texttt{(x)}. For
curly braces we enter \texttt{\{x\}}. And, lastly, for square brackets
we use \texttt{[x]}. These produce $(x)$, $\{x\}$, and $[x]$,
respectively. In longer expressions, though, the results can look
funny.
In
\[
[\sum_{x=1}^{3} \{ (\int_{0}^{x} (e^{y}~dy))\}],
\]
the size of the delimiters is off. However, in
\[
\left[\sum_{x=1}^{3} \left\{ \left(\int_{0}^{x} (e^{y}~dy)\right)\right\}\right],
\]
the problems seem to be gone. If we type \texttt{\textbackslash
  left(}, \texttt{\textbackslash left\textbackslash\{}, or
\texttt{\textbackslash left[} and close the expression off with the
approprate \texttt{\textbackslash right)}, \texttt{\textbackslash
  right\textbackslash\}}, or \texttt{\textbackslash right]}, the sizes
are determined by the size of the internal expression as opposed to
being constant.
\section{Accents}

Often times we need to modify slightly a character it denote that the
construct has slightly changed. For example, we want
\begin{itemize}
\item not the true parameter $\beta$ but the estimate $\hat{\beta}$,
\item not the random variable $y$, but the sample mean $\bar{y}$, or
\item not the singleton $x$, but a vector $\vec{x}$.
\end{itemize}

A partial list of these accents is below.
\begin{multicols}{3}
\begin{itemize}
\item[$\acute{a}$] \texttt{\textbackslash acute\{a\}}
\item[$\bar{b}$] \texttt{\textbackslash bar\{b\}}
\item[$\breve{c}$] \texttt{\textbackslash breve\{c\}}
\item[$\check{d}$] \texttt{\textbackslash check\{d\}}
\item[$\dot{e}$] \texttt{\textbackslash dot\{e\}}
\item[$\ddot{f}$] \texttt{\textbackslash ddot\{f\}}
\item[$\grave{g}$] \texttt{\textbackslash grave\{g\}}
\item[$\hat{h}$] \texttt{\textbackslash hat\{h\}}
\item[$\tilde{\imath}$] \texttt{\textbackslash tilde\{\textbackslash imath\}}
\item[$\vec{\jmath}$] \texttt{\textbackslash vec\{\textbackslash jmath\}}
\item[$\widehat{xyz}$] \texttt{\textbackslash widehat\{\textbackslash xyz\}}
\item[$\widetilde{xyz}$] \texttt{\textbackslash widetilde\{\textbackslash xyz\}}

\end{itemize}
\end{multicols}

\section{Assignment!}

\subsection*{Questions}
Write \LaTeX{} code for the following expressions. Confirm that your
compiled document matches these.
\begin{enumerate}
\item
\[
\neg\left(\left(T \vee F\right) \wedge F\right) \vee \left( T \right)
\]

\item 
\[
e^{ix} = \cos x + i \sin x
\]

\item
\[
y_i = \beta_0 + \beta_1 x_i + \beta_2 x_i^2 + \epsilon_i
\]

\item
\[
\left(\sum_{x=1}^{\infty} (1/x) \right) - \left(\sum_{x=1}^{\infty}
  (1/x)^2 \right)
\]
\end{enumerate}

\subsection*{Answers}

\begin{enumerate}

\item ~\\
  \begin{turn}{180}
  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}

\[
\neg\left(\left(T \vee F\right) 
\wedge F\right) \vee \left( T \right)
\]

\end{verbatim}
    \end{minipage}
  }
\end{turn}

\item ~\\
  \begin{turn}{180}
  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}

\[
e^{ix} = \cos x + i \sin x
\]

\end{verbatim}
    \end{minipage}
  }
\end{turn}

\item ~\\
  \begin{turn}{180}
  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}

\[
y_i = \beta_0 + \beta_1 x_i + 
\beta_2 x_i^2 + \epsilon_i
\]

\end{verbatim}
    \end{minipage}
  }
\end{turn}

\item ~\\
  \begin{turn}{180}
  \ovalbox{
    \begin{minipage}{\linewidth}
\begin{verbatim}

\[
\left(\sum_{x=1}^{\infty} (1/x) \right) - 
\left(\sum_{x=1}^{\infty}
  (1/x)^2 \right)
\]

\end{verbatim}
    \end{minipage}
  }
\end{turn}

\end{enumerate}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 